\LaTeXe\ guide for the Journal of Functional Programming

### 2019-05-02 papers/jfp/8-engineering.tex
Address Simon's feedback on S8.1

Another approach is to include failures into the computation context \hs{f}. For
example, we can choose \hs{f}~\hs{=}~\hs{Maybe}:

@@

\noindent
Since tasks are polymorphic in \hs{f}, we can directly run them with the
callback \hs{fetchA1A2}. For example, the task \cmd{B1 = A1 + A2} from
\hs{sprsh1} in~\S\ref{sec-task} returns \hs{Just}~\hs{30}, whereas the task
\cmd{B2 = B1 * 2} returns \hs{Nothing} because \hs{fetchA1A2} fails on \cmd{B1}.

This approach is convenient if \emph{tasks are not aware of failures}, e.g. we
can model Excel formulas as pure arithmetic functions, and introduce failures
``for free'' if/when needed by instantiating \hs{Tasks} with an appropriate
\hs{f}.

@@

\noindent
With this approach we can implement a build system that accepts
\hs{Tasks}~\hs{MonadFail}~\hs{k}~\hs{v} and handles errors by aborting the
build early and returning
\hs{Either}~\hs{String}~\hs{(Store}~\hs{i}~\hs{k}~\hs{v)} instead of just
\hs{Store}~\hs{i}~\hs{k}~\hs{v} as in~\S\ref{sec-general-build}. One possible
implementation is based on introducing an extra \hs{Either}~\hs{String} layer
into the monad stack; while fairly direct it is also tedious, and hence we omit
it.

>>>>

Another approach is to include failures into the computation context \hs{f}.
Recall from~\S\ref{sec-why-polymorphism} that we require tasks to be polymorphic
in~\hs{f}, and we can therefore choose to execute a \hs{Task} in an~\hs{f} with
failures, the simplest example being \hs{f}~\hs{=}~\hs{Maybe}. Below we define
a callback \hs{fetchA1A2} that returns \hs{Just}~\hs{value} for keys \cmd{A1}
and \cmd{A2} but fails with \hs{Nothing} on all other keys:

@@

\noindent
We can directly run any \hs{Task} with this callback. For example, the task
\cmd{B1 = A1 + A2} from \hs{sprsh1} in~\S\ref{sec-task} returns
\hs{Just}~\hs{30}, whereas the task \cmd{B2 = B1 * 2} returns \hs{Nothing}
because \hs{fetchA1A2} fails on \cmd{B1}.

This approach is convenient if \emph{tasks are not aware of failures}, e.g. we
can model \Excel formulas as pure arithmetic functions, and introduce failures
``for free'' if/when needed by instantiating \hs{Tasks} with an appropriate
\hs{f}. In a real system this \hs{f} would be more complex than just \hs{Maybe},
for example \hs{MaybeT}~\hs{(State}~\hs{Spreadsheet)}, thus allowing us to
combine failures with access to \Excel's spreadsheet state by using the
\hs{MaybeT} monad transformer~\cite{liang1995monad}.

@@

\noindent
With this approach we can implement a build system that accepts
\hs{Tasks}~\hs{MonadFail}~\hs{k}~\hs{v} and handles errors by aborting the
build early and returning
\hs{Either}~\hs{String}~\hs{(Store}~\hs{i}~\hs{k}~\hs{v)} instead of just
\hs{Store}~\hs{i}~\hs{k}~\hs{v} as in our \hs{Build}
abstraction~\S\ref{sec-general-build}. One possible implementation of such a
failure-handling build system is based on adding an extra
\hs{Either}~\hs{String} layer into the monad stack, e.g. augmenting the monad
\hs{State}~\hs{(Store}~\hs{i}~\hs{k}~\hs{v)} used by the schedulers
in~\S\ref{sec-implementations} with exceptions. We omit the actual
implementation: while fairly direct it is also tedious due to additional
wrapping and unwrapping.
